The Birth of a Galaxy: Primordial Metal Enrichment and Stellar Populations

### 2011-11-01 ApJ/ms.tex
Cleaning up star formation and feedback sections to be more clear.

\begin{figure*}
\epsscale{1.15}
\plotone{f1}
\epsscale{1}
  \caption{\label{fig:evo-mosaic} Evolution of the entire simulation
    volume ($L_{\rm box} = 1$ Mpc) at redshifts 15, 12, 10, 8, and 7
    (left to right).  Pictured here are the density-weighted
    projections of density (top), temperature (middle), and
    metallicity (bottom).  Note how the stellar radiative feedback
    from low-mass galaxies reionize the majority of the volume.  The
    metallicity projections are a composite image of metals
    originating from Pop II (blue) and III (red) stars with magenta
    indicating a mixture of the two.}
\end{figure*}

@@

@@

For Pop III stars, we use the same SF model as \citet{Wise08_Gal}
where each star particle represents a single star, forming at an
overdensity of $5 \times 10^5$ ($\sim 10^3\; \cubecm$ at $z=10$).  If
multiple cells meet the star particle formation criteria within 1 pc,
we form one Pop III star particle with the center of mass of these
flagged cells to ensure that massive star is created per metal-free
molecular cloud.  Instead of using a fixed stellar mass like in our
previous work, we randomly sample from an IMF with a functional form
of
%
\begin{equation}
\label{eqn:imf}
f(\log M) dM = M^{-1.3} \exp\left[-\left(\frac{M_{\rm
      char}}{M}\right)^{1.6}\right] dM
\end{equation}
to determine the stellar mass.  Above $M_{\rm char}$, it behaves as a
Salpeter IMF but is exponentially cutoff below that mass
\citep{Chabrier03, Clark09}.  For reproducibility, we record the
number of times the random number generator (Mersenne twister;
\citet{MTwister}) has been called for use when restarting the
simulations.  We stress that the Pop III IMF is still highly
uncertain.  In Sec. \ref{sec:discuss}, we discuss possible
dependencies of our results to changes in the IMF.

@@

We treat Pop II star formation with the same prescription as
\citet{Wise09}, which is similar to the Pop III prescription but
without the \hh~fraction floor that is removed because the
metal-enriched gas can efficiently cool even in the presence of a
strong UV radiation field \citep[e.g.][]{Safranek10}.  This method is
effectively a modified version of the \citet{Cen92} method but
accounts for the fact that the molecular clouds are resolved.  Unlike
Pop III star particles that represent individual stars, Pop II star
particles represent a star cluster of some total mass and an assumed
normal IMF.  In the simulation, we define the molecular cloud as a
sphere, centered on the most dense cell, with a dynamical time $t_{\rm
  dyn} = 3$ Myr (corresponding to an average density $\bar{\rho}_{\rm
  MC} \simeq 1000\mu \cubecm$) and a radius $R_{\rm MC}$, where $\mu$
is the mean molecular weight.  In this sphere, a fraction $c_\star =
0.07 f_{\rm cold}$ of the cold gas ($T < 10^3$ K) is converted into a
star particle with mass $m_\star = c_\star (4\pi/3) \bar{\rho}_{\rm
  MC} R_{\rm MC}^3$.  We then replace the sphere with a uniform
density $\rho_{\rm MC} = (1 - c_\star) / (Gt_{\rm dyn}^2)$ and
temperature $T = 10^4$ K, which approximates the initial stages of an
\ion{H}{2} region.  Our choice of $c_\star$ agrees with the theory
proposed in \citet{Krumholz05} that are supported by observations.
Observations of star-forming regions show that the dynamical time is
approximately 700 kyr and that SF is extended over several dynamical
times \citep[e.g.][]{Tan06}, which motivates our choice of $t_{\rm
  dyn}$ of our definition of molecular clouds in our SF prescription.

Similar to the Pop III treatment, we merge any newly created particles
within a radius of 10 pc.  The new star particle acquires the center
of mass of the merged particles.  During the merging, the gas is
unaltered besides the initial gas accretion from the grid to the star
particles.  We set the minimum mass of a star particle to $m_{\rm
  \star, min} = 1000 \Ms$.  If the initial mass does not exceed
$m_{\rm \star, min}$, the star particle does not provide any feedback
and continues to accrete until it reaches $m_{\rm \star, min}$.  The
star particles live for 20 Myr, the maximum lifetime of an OB star.
These stars generate the majority of the ionizing radiation and SNe
feedback in stellar clusters, thus we ignore any feedback from lower
mass stars.

@@

We use mass-dependent hydrogen ionizing and LW photon luminosities and
lifetimes of the Pop III stars from \citet{Schaerer02}.  We use a
mass-independent photon energy $E_{\rm ph} = 29.6$ eV, appropriate for
the near-constant $10^5$ K surface temperatures of Pop III stars.
They die as Type II SNe if $11 \le M_\star/\Ms \le 40$
\Ms~\citep{Woosley95} and as PISNe if they are in the mass range
140--260 \Ms~\citep{Heger03}, where $M_\star$ is the stellar mass.
For normal Type II SNe between 11--20 \Ms, we use an explosion energy
of \tento{51}~erg and a linear fit to the metal ejecta mass calculated
in \citet{Nomoto06},
%
\begin{equation}
  \label{eqn:typeii}
  M_{\rm Z}/\Ms = 0.1077 + 0.3383 \times (M_\star/\Ms - 11).
\end{equation}
%
We model the SNe of stars with $20 \le M_\star/\Ms \le 40$ as
hypernova with the energies and ejecta masses also taken from
\citeauthor{Nomoto06}, linearly interpolating their results to
$M_\star$.  For PISNe, we use the explosion energy from
\citet{Heger02}, where we fit the following function to their models,
%
\begin{equation}
  \label{eqn:pisn}
  E_{\rm PISN} = 10^{51} \times [5.0 + 1.304 (M_{\rm He}/\Ms - 64)] \; \mathrm{erg},
\end{equation}
%
where $M_{\rm He} = (13/24) \times (M_\star - 20) \Ms$ is the helium
core (and equivalently the metal ejecta) mass and $M_\star$ is the
stellar mass.  If the stellar mass is outside of these ranges, then an
inert, collisionless black hole (BH) particle is created.

The blastwave is modeled by injecting the explosion energy and ejecta
mass into a sphere of 10 pc, smoothed at its surface to improve
numerical stability \citep{Wise08_Gal}.  Because we resolve the
blastwave relatively well with several cells across at its
initialization, the thermal energy is converted into kinetic energy
and agrees with the Sedov-Taylor solution \citep[e.g.][]{Greif07}.

The Pop II stars emit 6000 hydrogen ionizing photons per baryon over
their lifetime and $E_{\rm ph} = 21.6$ eV, appropriate for a [Z/H] =
$-1.3$ population.  We note that lower-metallicity stars generate up
to 60\% more ionizing photons and might be underestimating the
radiative feedback \citep{Schaerer03}.  Nonetheless this study
provides an excellent first insight in the transition to Pop II SF, as
the metal enrichment is the key ingredient here.  For SN feedback,
these stars generate $6.8 \times 10^{48}$ erg s$^{-1}$ $\Ms^{-1}$
after living for 4 Myr, which is injected into spheres with a radius
of 10 pc.  However if the resolution of the grid is less than 10/3 pc,
we distribute the energy into a $3^3$ cube surrounding the star
particle.  In the same region, star particles also return ejected
material with mass
%
\begin{equation}
  \label{eqn:SNmass}
  \Delta m_{\rm ej} = \frac{0.25 \; \Delta t \; M_\star} 
         {t_0 - 4\;\mathrm{Myr}}
\end{equation}
back to the grid at every timestep on the finest AMR level.  Here
$t_0$ = 20 Myr is the lifetime of the star particle.  This ejected gas
has solar metallicity $Z = 0.02$, resulting in a total metal yield $y
= 0.005$.  In addition to numerical stability, energy distribution
across several cells instead of a single cell has been shown to match
cosmic SFRs better in galaxy simulations at lower redshifts
\citep{Smith11}.

@@

\begin{figure}
\epsscale{1.15}
\plotone{f2}
\epsscale{1}
  \caption{\label{fig:halo-mosaic} Density-weighted projections of
    density (left), temperature (center), and metallicity (right) of
    the two selected halos at $z=7$.  The field of view is 5 proper
    kpc, and the circles have a radius of $r_{200}$.  The composite
    metallicity images are constructed in the same fashion as
    Fig. \ref{fig:evo-mosaic}.}
\end{figure}

@@

\begin{figure}
\epsscale{1.15}
\plotone{f3}
\epsscale{1}
\caption{\label{fig:sfr} Cosmic star formation rates for Pop II
  (solid) and III (dashed).  Note that the Pop III rates are nearly
  constant from redshift 12 to 7.  Pristine halos still survive in
  underdense regions, supporting late-time metal-free star formation.}
\end{figure}

@@

@@

\begin{figure*}
\epsscale{1.15}
\plotone{f4}
\epsscale{1}
\caption{\label{fig:evo} (a) Evolution of the total halo mass (top),
  stellar mass (middle), and gas fraction (bottom) of the quiet
  (dashed) and intense (solid) halos.  In the top panel, the filtering
  halo mass is plotted as the thick black line (overlapping with the
  quiet halo after $z=9$.  (b) Mass-weighted stellar metallicities and
  gas metallicities enriched by Pop II and Pop III SNe of the intense
  (top) and quiet (bottom) halos.}
\end{figure*}

@@

\begin{figure}
\epsscale{1.15}
\plotone{f5}
\epsscale{1}
\caption{\label{fig:z3} Average halo metallicity from Pop III SNe as a
  function of total halo mass.  Inhomogeneous metal enrichment causes
  the large spread in metallicity at low masses.  At $M_{\rm halo} > 3
  \times 10^7 \Ms$, all halos are enriched by Pop III SNe above [Z/H]
  $>$ --4.  The y-axis error bars show the 25\% and 75\% percentile
  values because the underlying distribution is not Gaussian.}
\end{figure}

@@

\begin{figure}
\epsscale{1.15}
\plotone{f7}
\epsscale{1}
\caption{\label{fig:massfn} Total Pop II luminosities of halos versus
  total halo mass with constant mass-to-light ratios overplotted at
  redshift 7.  The halos with $M \gsim 10^8 \Ms$ form stars
  efficiently because they can cool through atomic line cooling,
  whereas the lower mass galaxies rely on molecular hydrogen cooling
  only.  The low M/L ratios and spread across two orders of magnitude
  are similar to ones observed in nearby ultra faint dwarf galaxies.}
\end{figure}

@@

%\clearpage
%\input biblio.tex
\bibliography{ms}
\end{document}

>>>>

@@

\subsubsection{Population III star formation}

@@

For Pop III stars, we use the same SF model as \citet{Wise08_Gal}
where each star particle represents a single star.  In this model, a
star particle forms when a cell has all of the following criteria:
%
\begin{enumerate}
\item an overdensity of $5 \times 10^5$ ($\sim 10^3\; \cubecm$ at
  $z=10$),
\item a converging gas flow ($\nabla \cdot \mathbf{v}_{\rm gas} < 0$),
  and
\item a molecular hydrogen fraction $f_{\rm H2} > 5 \times 10^{-4}$.
\end{enumerate}
%
These physical conditions are typical of collapsing metal-free
molecular clouds $\sim 10$ Myr before the birth of a Pop III
main-sequence star \citep{ABN02, OShea07a}.  This prescription is
similar to the \citet{Cen92} method but removes the Jeans unstable
($M_{\rm gas} > M_J$) and cooling timescale ($t_{\rm cool} < t_{\rm
  dyn}$) requirements.  We do not consider the former criterion
because it is not applicable to simulations that resolve the Jeans
length at all times.  The molecular hydrogen fraction requirement
effectively constrains star formation to cooling molecular clouds,
where the \hh~formation rate is significantly larger than the
dissociation rate from a Lyman-Werner radiation field.

If multiple cells meet the star particle formation criteria within 1
pc, we form one Pop III star particle with the center of mass of these
flagged cells to ensure that one massive star is created per
metal-free molecular cloud.  Instead of using a fixed stellar mass
like in our previous work, we randomly sample from an IMF with a
functional form of
%
\begin{equation}
\label{eqn:imf}
f(\log M) dM = M^{-1.3} \exp\left[-\left(\frac{M_{\rm
      char}}{M}\right)^{1.6}\right] dM
\end{equation}
%
to determine the stellar mass.  Above $M_{\rm char}$, it behaves as a
Salpeter IMF but is exponentially cutoff below that mass
\citep{Chabrier03, Clark09}.  For reproducibility, we record the
number of times the random number generator (Mersenne twister;
\citet{MTwister}) has been called for use when restarting the
simulations.  We stress that the Pop III IMF is still highly
uncertain.  In Sec. \ref{sec:discuss}, we discuss possible
dependencies of our results to changes in the IMF.

After the star particle forms and its mass is randomly sampled from
the IMF, an equal amount of gas is removed from the computational grid
in a sphere that contains twice the stellar mass and is centered on
the star particle.  The star particle acquires the mass-weighted
velocity of the gas contained in this sphere.

@@

\subsubsection{Population II star formation}

We treat Pop II star formation with the same prescription as
\citet{Wise09}, which is similar to the Pop III prescription but
without the minimum \hh~fraction requirement.  This is removed because
the metal-enriched gas can efficiently cool even in the presence of a
strong UV radiation field \citep[e.g.][]{Safranek10}.  To ensure the
volume is cooling, we restrict star formation to gas with temperatures
$T < 1000$ K.  Unlike Pop III star particles that represent individual
stars, Pop II star particles represent a star cluster of some total
mass and an assumed normal IMF.

Once a cell meets these criteria, the prescription searches outward
with increasing radius for the boundary of the molecular cloud,
centered on the most dense cell.  Here, we define a molecular cloud as
a sphere with a dynamical time $t_{\rm dyn} = 3$ Myr (corresponding to
an average density $\bar{\rho}_{\rm MC} \simeq 1000\mu\; \cubecm$) and
a radius $R_{\rm MC}$, where $\mu$ is the mean molecular weight.  Once
the sphere radius is found, a fraction $c_\star = 0.07 f_{\rm cold}$
of the cold gas ($T < 10^3$ K) inside the sphere is converted into a
star particle with mass $m_\star = c_\star (4\pi/3) \bar{\rho}_{\rm
  MC} R_{\rm MC}^3$.  This treatment of cold gas accretion is similar
to a star formation model with a multi-phase subgrid model
\citep{Springel03b} but is employed in a simulation that can resolve
the multi-phase interstellar stellar medium (ISM).

After the star particle is created, we replace the sphere with a
uniform density $\rho_{\rm MC} = (1 - c_\star) / (Gt_{\rm dyn}^2)$ and
temperature $T = 10^4$ K, which approximates the initial stages of an
\ion{H}{2} region.  Our choice of $c_\star$ agrees with the theory
proposed in \citet{Krumholz05} that are supported by observations.
Observations of star-forming regions show that the dynamical time is
approximately 700 kyr and that SF is extended over several dynamical
times \citep[e.g.][]{Tan06}, which motivates our choice of $t_{\rm
  dyn}$ of our definition of molecular clouds in our SF prescription.

Similar to the Pop III treatment, we merge any newly created particles
within a radius of 10 pc.  The new star particle acquires the center
of mass of the merged particles.  During the merging, the gas is
unaltered besides the initial gas accretion from the grid to the star
particles.  We set the minimum mass of a star particle to $m_{\rm
  \star, min} = 1000 \Ms$.  If the initial mass does not exceed
$m_{\rm \star, min}$, the star particle does not provide any feedback
and continues to accrete until it reaches $m_{\rm \star, min}$.

@@

\subsubsection{Population III stellar feedback}

We use mass-dependent hydrogen ionizing and LW photon luminosities and
lifetimes of the Pop III stars from \citet{Schaerer02}.  We use a
mass-independent photon energy $E_{\rm ph} = 29.6$ eV, appropriate for
the near-constant $10^5$ K surface temperatures of Pop III stars.
They die as Type II SNe if $11 \le M_\star/\Ms \le 40$
\Ms~\citep{Woosley95} and as PISNe if they are in the mass range
140--260 \Ms~\citep{Heger03}, where $M_\star$ is the stellar mass.
For normal Type II SNe between 11--20 \Ms, we use an explosion energy
of \tento{51}~erg and a linear fit to the metal ejecta mass calculated
in \citet{Nomoto06},
%
\begin{equation}
  \label{eqn:typeii}
  M_{\rm Z}/\Ms = 0.1077 + 0.3383 \times (M_\star/\Ms - 11).
\end{equation}
%
We model the SNe of stars with $20 \le M_\star/\Ms \le 40$ as
hypernova with the energies and ejecta masses also taken from
\citeauthor{Nomoto06}, linearly interpolating their results to
$M_\star$.  For PISNe, we use the explosion energy from
\citet{Heger02}, where we fit the following function to their models,
%
\begin{equation}
  \label{eqn:pisn}
  E_{\rm PISN} = 10^{51} \times [5.0 + 1.304 (M_{\rm He}/\Ms - 64)] \;
  \mathrm{erg},
\end{equation}
%
where $M_{\rm He} = (13/24) \times (M_\star - 20) \Ms$ is the helium
core (and equivalently the metal ejecta) mass and $M_\star$ is the
stellar mass.  If the stellar mass is outside of these ranges, then an
inert, collisionless black hole (BH) particle is created.  The
blastwave is modeled by injecting the explosion energy and ejecta mass
into a sphere of 10 pc, smoothed at its surface to improve numerical
stability \citep{Wise08_Gal}.  Because we resolve the blastwave
relatively well with several cells across at its initialization, the
thermal energy is converted into kinetic energy and agrees with the
Sedov-Taylor solution \citep[e.g.][]{Greif07}.

\subsubsection{Population II star formation}

The Pop II stars emit 6000 hydrogen ionizing photons per baryon over
their lifetime and $E_{\rm ph} = 21.6$ eV, appropriate for a [Z/H] =
$-1.3$ population.  We note that lower-metallicity stars generate up
to 60\% more ionizing photons and might be underestimating the
radiative feedback \citep{Schaerer03}.  Nonetheless this study
provides an excellent first insight in the transition to Pop II SF, as
the metal enrichment is the key ingredient here.  The star particles
live for 20 Myr, the maximum lifetime of an OB star.  These stars
generate the majority of the ionizing radiation and SNe feedback in
stellar clusters, thus we ignore any feedback from lower mass stars.

For SN feedback, these stars generate $6.8 \times 10^{48}$ erg
s$^{-1}$ $\Ms^{-1}$ after living for 4 Myr, which is injected into
spheres with a radius of 10 pc.  However if the resolution of the grid
is less than 10/3 pc, we distribute the energy into a $3^3$ cube
surrounding the star particle.  In the same region, star particles
also return ejected material with mass
%
\begin{equation}
  \label{eqn:SNmass}
  \Delta m_{\rm ej} = \frac{0.25 \; \Delta t \; M_\star} 
         {t_0 - 4\;\mathrm{Myr}}
\end{equation}
back to the grid at every timestep on the finest AMR level.  Here
$t_0$ = 20 Myr is the lifetime of the star particle.  This ejected gas
has solar metallicity $Z = 0.02$, resulting in a total metal yield $y
= 0.005$.  In addition to numerical stability, energy distribution
across several cells instead of a single cell has been shown to match
cosmic SFRs better in galaxy simulations at lower redshifts
\citep{Smith11}.

@@

\begin{figure*}
%\epsscale{1.15}
\plotone{f1}
\epsscale{1}
  \caption{\label{fig:evo-mosaic} Evolution of the entire simulation
    volume ($L_{\rm box} = 1$ Mpc) at redshifts 15, 12, 10, 8, and 7
    (left to right).  Pictured here are the density-weighted
    projections of density (top), temperature (middle), and
    metallicity (bottom).  Note how the stellar radiative feedback
    from low-mass galaxies reionize the majority of the volume.  The
    metallicity projections are a composite image of metals
    originating from Pop II (red) and III (blue) stars with magenta
    indicating a mixture of the two.}
\end{figure*}

@@

\begin{figure}
%\epsscale{1.15}
\plotone{f2}
\epsscale{1}
  \caption{\label{fig:halo-mosaic} Density-weighted projections of
    density (left), temperature (center), and metallicity (right) of
    the two selected halos at $z=7$.  The field of view is 5 proper
    kpc, and the circles have a radius of $r_{200}$.  The composite
    metallicity images are constructed in the same fashion as
    Fig. \ref{fig:evo-mosaic}.}
\end{figure}

@@

\begin{figure}
%\epsscale{1.15}
\plotone{f3}
\epsscale{1}
\caption{\label{fig:sfr} Cosmic star formation rates for Pop II
  (solid) and III (dashed).  Note that the Pop III rates are nearly
  constant from redshift 12 to 7.  Pristine halos still survive in
  underdense regions, supporting late-time metal-free star formation.}
\end{figure}

@@

\begin{figure*}
\epsscale{1}
\plotone{f4}
\epsscale{1}
\caption{\label{fig:evo} (a) Evolution of the total halo mass (top),
  stellar mass (middle), and gas fraction (bottom) of the quiet
  (dashed) and intense (solid) halos.  In the top panel, the filtering
  halo mass is plotted as the thick black line (overlapping with the
  quiet halo after $z=9$.  (b) Mass-weighted stellar metallicities and
  gas metallicities enriched by Pop II and Pop III SNe of the intense
  (top) and quiet (bottom) halos.}
\end{figure*}

@@

\begin{figure}
\epsscale{1}
\plotone{f5}
\epsscale{1}
\caption{\label{fig:z3} Average halo metallicity from Pop III SNe as a
  function of total halo mass.  Inhomogeneous metal enrichment causes
  the large spread in metallicity at low masses.  At $M_{\rm halo} > 3
  \times 10^7 \Ms$, all halos are enriched by Pop III SNe above [Z/H]
  $>$ --4.  The y-axis error bars show the 25\% and 75\% percentile
  values because the underlying distribution is not Gaussian.}
\end{figure}

@@

\begin{figure}
\epsscale{1}
\plotone{f7}
\epsscale{1}
\caption{\label{fig:massfn} Total Pop II luminosities of halos versus
  total halo mass with constant mass-to-light ratios overplotted at
  redshift 7.  The halos with $M \gsim 10^8 \Ms$ form stars
  efficiently because they can cool through atomic line cooling,
  whereas the lower mass galaxies rely on molecular hydrogen cooling
  only.  The low M/L ratios and spread across two orders of magnitude
  are similar to ones observed in nearby ultra faint dwarf galaxies.}
\end{figure}

@@

%\clearpage
\bibliography{ms}

\end{document}
